Deciding to Disclose: A Decision Theoretic Agent Model of Pregnancy and Alcohol Misuse

### 2014-10-29 Abstract.tex
Ammendments

We draw together methodologies from game theory, agent based modelling, and decision theory to explore the process of decision making around disclosure. This is framed in the context of pregnant women disclosing their drinking behaviour to their midwives.

@@

The primary purpose it to demonstrate the potential utility of an approach which it is hoped goes some way towards addressing concerns about the ad hoc character of \ac{ABM}, by providing a strong theoretical grounding for the reasoning processes of individual agents. To this end we hope to show that these simple rules, operating in an inescapably artificial scenario are nonetheless capable of producing trends from the literature.
We also seek to demonstrate the significance of precisely how the decision making process is formulated, by contrasting four distinct decision rules against one another and exploring a simple form of information sharing, supported by the use of statistical emulators for a full exploration of the parameter space.

@@

We employ game theory to define a signalling game representative of a scenario where pregnant women decide how far to disclose their drinking behaviours to their midwives, and midwives employ the information provided to decide whether a costly referral should be made. This game is then recast as two games taking play against nature, to permit the use of a decision theoretic approach where both classes of agent use simple rules to decide their moves.
Four decision rules are explored - a lexicographic heuristic which considers only the link between moves and payoffs, a Bayesian risk minimisation agent that uses the same information, a more complex Bayesian risk minimiser, and a \ac{CPT} type.

Using a simulator we have developed in Python, we recreate two key qualitative trends described in the Midwifery literature for all the decision models, and investigate the impact of introducing a simple form of information sharing within agent groups.
Finally a global sensitivity analysis using \acp{GEM} was conducted, to compare the response surfaces of the different decision rules in the game.

@@

We note that the scenario omits the overwhelming complexity of the reality, and is presented largely in the spirit of a convenient demonstration of the methodology. Clearly a domain where there is sufficient data to permit a more comprehensive approach to validation of model outcomes is desirable, and will form the basis of our future work.

>>>>

We draw together methodologies from game theory, agent based modelling, decision theory, and uncertainty analysis to explore the process of decision making in the context of pregnant women disclosing their drinking behaviour to their midwives.

@@

The primary purpose of this work is to demonstrate the significance of precisely how the decision making process is formulated, by contrasting four distinct decision rules against one another and exploring a simple form of information sharing, supported by the use of statistical emulators for a full exploration of the parameter space.
We also demonstrate the utility of an approach which it is hoped goes some way towards addressing concerns about the ad hoc character of \ac{ABM}, by providing a strong theoretical grounding for the reasoning processes of individual agents. To this end we hope to show that these simple rules, operating in an inescapably artificial scenario are nonetheless capable of reproducing trends from the literature.

@@

We employ game theory to define a signalling game representative of a scenario where pregnant women decide the extent to which they disclose their drinking behaviours to their midwives, and midwives employ the information provided to decide whether to refer their patients for costly specialist treatment. This game is then recast as two games played against nature, to permit the use of a decision theoretic approach where both classes of agent use simple rules to decide their moves.
Four decision rules are explored - a lexicographic heuristic which considers only the link between moves and payoffs, a Bayesian risk minimisation agent that uses the same information, a more complex Bayesian risk minimiser with full access to the structure of the decision problem, and a \ac{CPT} rule.

In simulation, we recreate two key qualitative trends described in the Midwifery literature for all the decision models, and investigate the impact of introducing a simple form of information sharing within agent groups.
Finally a global sensitivity analysis using \acp{GEM} was conducted, to compare the response surfaces of the different decision rules in the game.

@@

We note that the model omits the overwhelming complexity of the real disclosure scenario, and is presented largely in the spirit of a convenient demonstration of the methodology. Clearly a domain where there is sufficient data to permit a more comprehensive approach to validation of model outcomes is desirable, and will form the basis of our future work.

### 2014-10-29 Background.tex
Ammendments

This basic form has been widely applied, with substantial interest in what conditions permit honest signalling as Nash equilibria or \ac{ESS}. \citet{Grafen1990}, following from a suggestion by \citet{Zahavi1975}, proposed that if signals intended to indicate mate quality exacted a cost on the signaller (e.g. peacock tail feathers), then honest signalling would constitute an \ac{ESS}. Similar results have also been demonstrated in a game of job market signalling, where signal cost was differentiated by type \citep{Spence1973}. 
Costly signalling has also been suggested as an explaination of behaviour that at first gasp appears counter intuitive, for example \citet{Godfray1991} applied the idea to the food solicitation behaviour of chicks, where a stronger signal carries a risk of being eaten. Moving beyond animal behaviour, \citet{Sosis2003} considered the implications if ritual behaviour, in the context of religion, represented a costly signal, an idea subsequently extended by \citet{Henrich2009} to include cultural transmission, and \citet{Wildman2011} to introduce group differentiation.

@@

An alternative normative model of decision making is Bayesian decision theory, proposed by \citet{Robbins1964}, which is essentially the application of Bayesian style probabilities to the expected utility model. This allows probabilities used in reasoning to be subjective, which may allow for a better account of decisions from experience (see \citet{Hau2008,Hertwig2004} for results elucidating the distinction, and comparing the performance of several non-Bayesian models). This model has seen notable successes in practical problems \citep{McNamara1980,Survey2003,Kristensen1997}, but suggestions by several authors  that it could constitute an effective (top-down) model of learning \citep{Tenenbaum2006,Griffiths2010}, or induction \citep{Gallistel2012} in the brain have attracted substantil criticism, e.g. \citet{Bowers2012}, and \citet{Miller2012} responding to \citeauthor{Tenenbaum2006,Griffiths2010}, and \citeauthor{Gallistel2012} respectively.

>>>>

This basic form has been widely applied, with substantial interest in what conditions permit honest signalling as Nash equilibria or \acp{ESS}. \citet{Grafen1990}, following from a suggestion by \citet{Zahavi1975}, proposed that if signals intended to indicate mate quality exacted a cost on the signaller (e.g. peacock tail feathers), then honest signalling would constitute an \ac{ESS}. Similar results have also been demonstrated in a game of job market signalling, where signal cost was differentiated by type \citep{Spence1973}. 
Costly signalling has also been suggested as an explaination of behaviour that at first gasp appears counter intuitive, for example \citet{Godfray1991} applied the idea to the food solicitation behaviour of chicks, where a stronger signal carries a risk of being eaten. Moving beyond animal behaviour, \citet{Sosis2003} considered the implications if ritual behaviour, in the context of religion, represented a costly signal, an idea subsequently extended by \citet{Henrich2009} to include cultural transmission, and \citet{Wildman2011} to introduce group differentiation.

@@

An alternative normative model of decision making is Bayesian decision theory, proposed by \citet{Robbins1964}, which is essentially the application of Bayesian style probabilities to the expected utility model. This allows probabilities used in reasoning to be subjective, which may allow for a better account of decisions from experience (see \citet{Hau2008,Hertwig2004} for results elucidating the distinction, and comparing the performance of several non-Bayesian models). This model has seen notable successes in practical problems \citep{McNamara1980,Survey2003,Kristensen1997}, but suggestions by several authors  that it could constitute an effective (top-down) model of learning \citep{Tenenbaum2006,Griffiths2010}, or induction \citep{Gallistel2012} in the brain have attracted substantil criticism, e.g. \citet{Bowers2012} responding to \citeauthor{Tenenbaum2006}, and \citeauthor{Griffiths2010}; and \citet{Miller2012} addressing \citeauthor{Gallistel2012}.

### 2014-10-29 Conclusion.tex
Ammendments

As noted in section \ref{sub:sensitivity}, the impact of information sharing between midwives is surprisingly minimal, where it might be expected to play a more significant role in reality. A possible explanation for this lies in the implementation, which may place an excessive constraint on how much information midwives can share. The restriction to sharing only after a referral, together with the disparity in population sizes, and random allocation of appointments leads to midwives rarely having more than a single interaction with woman to pass on to their colleagues. In reality it might be anticipated that midwives would not withhold judgement so completely, and would concerns about particular women to their colleagues. This might be addressed by incorporating noisy type information \citep{Feltovich2002}, capturing the unintentional information transmitted during appointments, together with a relaxation of the assumptions about when information may be shared.  This also highlights an advantage of the \ac{BACCO} approach, in that it aids in diagnosing issues with model design by giving insight into parameters which are contributing innapropriately to variance in output.

>>>>

As noted in section \ref{sub:sensitivity}, the impact of information sharing between midwives is surprisingly minimal, where it might be expected to play a more significant role in reality. A possible explanation for this lies in the implementation, which may place an excessive constraint on how much information midwives can share. The restriction to sharing only after a referral, together with the disparity in population sizes, and random allocation of appointments leads to midwives rarely having more than a single interaction with woman to pass on to their colleagues. Furthermore, because midwives are only informed of the true type if a referral occurs, they have an inherent myopia, since until they have evidence of deception they will not refer, with said evidence difficult to obtain without a referral.

In reality it might be anticipated that midwives would not withhold judgement so completely, and would pass on concerns about particular women to their colleagues, or that particularly dramatic stories would persist and be passed. This might be addressed by incorporating noisy type information \citep{Feltovich2002}, capturing the unintentional information transmitted during appointments, together with a relaxation of the assumptions about when information may be shared and more sophisticated model of information transmission in general. This also highlights an advantage of the \ac{BACCO} approach, in that it aids in diagnosing issues with model design by giving insight into parameters which are contributing innapropriately to variance in output.

### 2014-10-29 Introduction.tex
Ammendments

The case in favour of \ac{ABM} as a general approach has been made numerously, and elegantly \citep{Silverman2011,Silverman2013,Resnick,gilbert1999simulation,epstein1994growing,Macy2002a,Axelrod1997}. As such we will not belabour the point, and instead turn to addressing some of the concerns expressed about the method. In this instance we focus on the perception of \ac{ABM} as ad hoc in nature, tending to be a reflection of the assumptions of the modeller rather than empirically or theoretically grounded \citep{Waldherr2013}. To ameliorate this concern, we extend previous work by \cite{GrayDissert} drawing on decision theory to produce simple rule based, learning, decision making agents and show that they are able to play a form of signalling game \citep{Kreps1987} with a basic form of intragroup information sharing. Four decision models of varying complexity, and behavioural plausibility are contrasted, by way of demonstrating the significance of the operationalisation of decision making in \ac{ABM}.

This is framed in the context of disclosure decisions, and a scenario examining drinking patterns in pregnant women which is presented in the spirit of a motivating example, rather than claimed as an accurate representation of reality. Alcohol consumption in the ante natal period is a significant issue in itself, and has been associated with many potentially negative consequences. For example, \citet{Andersen2012} report results from a large scale Danish cohort study suggesting that even low levels of consumption in early pregnancy increase the risk of spontaneous abortion, although \citet{Savitz2012} has suggested this may be attributable to a previously known link to absence of mornining sickness. Risk continues to the point of birth - \citet{Kesmodel2002} found a heightened risk to the infant - into childhood, with a metastudy by \citet{Latino-Martel2010} finding evidence on an increased risk of childhood \ac{AML} (although they suggest that the rarity of the condition is a limitation). Harm may also extend even further, and a review by \citet{Huizink2006} concluded that maternal alcohol consumption could be a contributing factor to \ac{ADHD}, and other learning impairments, but note methodological issues in a number of the papers. There is not, however, a clear consensus, with \citet{Gray2006} finding no evidence of harm below 1.5 UK units per day. In terms of official guidance, \ac{NICE} acknowledge that evidence of harm to the fetus is less than conclusive, but advise not drinking at all, or significant moderation \citep{NICE2010a}, with similar advice from the \cite{DepartmentofHealth2008}.

>>>>

The case in favour of \ac{ABM} as a general approach has been made numerously, and elegantly (e.g. \citet{Silverman2011,Silverman2013,Resnick,gilbert1999simulation,epstein1994growing,Macy2002a,Axelrod1997}, amongst others). As such we will not belabour the point, and instead turn to addressing some of the concerns expressed about the method. In this instance we focus on the perception of \ac{ABM} as ad hoc in nature, tending to be a reflection of the assumptions of the modeller rather than empirically or theoretically grounded \citep{Waldherr2013}. To ameliorate this concern, we extend previous work by \cite{GrayDissert} drawing on decision theory to produce simple rule based, learning, decision making agents and show that they are able to play a form of signalling game \citep{Kreps1987} with a basic form of intragroup information sharing. Four decision models of varying complexity, and behavioural plausibility are contrasted, by way of demonstrating the significance of the operationalisation of decision making in \ac{ABM}.

This is framed in the context of disclosure decisions, and a scenario examining drinking patterns in pregnant women which is presented in the spirit of a motivating example, rather than claimed as an accurate representation of reality. Alcohol consumption in the ante natal period is a significant issue in itself, and has been associated with many potentially negative consequences. For example, \citet{Andersen2012} report results from a large scale Danish cohort study suggesting that even low levels of consumption in early pregnancy increase the risk of spontaneous abortion, although \citet{Savitz2012} has suggested this may be attributable to a previously known link to absence of mornining sickness. Risk continues to the point of birth - \citet{Kesmodel2002} found a heightened risk to the infant - into childhood, with a metastudy by \citet{Latino-Martel2010} finding evidence on an increased risk of childhood \ac{AML}. Harm may also extend even further, and a review by \citet{Huizink2006} concluded that maternal alcohol consumption could be a contributing factor to \ac{ADHD}, and other learning impairments, but note methodological issues in a number of the papers. There is not, however, a clear consensus, with \citet{Gray2006} finding no evidence of harm below 1.5 UK units per day. In terms of official guidance, \ac{NICE} acknowledge that evidence of harm to the fetus is less than conclusive, but advise not drinking at all, or significant moderation \citep{NICE2010a}, with similar advice from the \cite{DepartmentofHealth2008}.

### 2014-10-29 Model.tex
Ammendments

Taken together, this leads to a game tree that is relatively complex even at the subgame level. Rather than attempt to solve for equilibria, agents treat this two player game as taking place against nature, along the lines of adversial risk analysis \citep{RiosInsua2009}. This effectively translates the game to a pair of decision problems (figure \ref{fig:decision_probs}), which agents attempt to resolve at each turn using a simple decision rule, given their prior beliefs and experience of play.

@@

As noted in section \ref{sub:the_game}, agents have perfect recall, and recognise individual opponents if they encounter them subsequently. While agents recall perfectly and make use of the new information for retrospective updates, all four agent models make decisions `as-if' they were always facing a new opponent.

@@

The lexicographic heuristic (algorithm \ref{alg:lexico}) follows the form of that used in \cite{Hau2008}, and assumes a simplified decision problem, as in figure \ref{fig:payoff_problem}, where an action is a choice between combined lotteries. Functionally, the heuristic maintains a count of the number of times that each action was followed by a payoff, and chooses the action which most commonly has the best payoff, i.e. one reason decision making. This approach requires minimal computation, and does not assume that \(u_{i}\) is static, or known.

@@

%Maybe cut this bit and make explanation richer?
Rather than the psychologically more interesting \ac{PT}, the \ac{CPT}
decision rule is used in this instance, because of the requirement
for women to evaluate more than two `prospects'.\footnote{A prospect in this instance is a paired outcome and probability, and the set of prospects
for an action hence define the outcome space.} \ac{CPT} introduces
the concept of a probability weighting function, which underweights
small probabilities, and overweights large ones in an effort to capture
the tendency of humans to treat high probability events as sure things,
and small probabilities as `never going to happen'. A number of different
weighting functions have been proposed, but in this instance the original
formulation by \citet{Tversky1992} is used. This distinguishes between
weighting for gains, and losses -

@@

Where $p$ is the unweighted probability, and $\gamma$ and $\delta$
are the weights for gain and loss probabilities respectively. Humans
have also been observed to value gains and losses differently, with
a loss being `worse' than the equivalent gain is `good'. This
entails a transformed value function -

@@

And $\alpha$, and $\beta$ are respectively the power of a gain,
and a loss, and \(x=u_{i}\). The \ac{CPT} value of outcome $x$ is $v(x)w^{+}(x)$
if $x\geq0$, and $v(x)w^{-}(x)$ otherwise. For an action the \ac{CPT}
value is the sum of the value of the prospects of that action, as
in the Bayesian risk model. The decision rule then requires the agent
to choose the action which maximises the prospect theory value.

>>>>

Taken together, this leads to a game tree that is relatively complex (figure \ref{fig:simple_tree} shows a simplified representation) even at the subgame level. Rather than attempt to solve for equilibria, agents treat this two player game as taking place against nature, along the lines of adversial risk analysis \citep{RiosInsua2009}. This effectively translates the game to a pair of decision problems (figure \ref{fig:decision_problems}), which agents attempt to resolve at each turn using a simple decision rule, given their prior beliefs and experience of play.

\begin{figure}[H]
\begin{adjustbox}{center}
\subfloat[Women]{\includegraphics[width=0.43\paperwidth]{figures/simple_game_tree}}
\end{adjustbox}
\caption{Simmplified game tree.}

\label{fig:simple_tree}
\end{figure}

@@

As noted in section \ref{sub:the_game}, agents have perfect recall, and midwives recognise women if they encounter them subsequently. While agents recall perfectly and make use of the new information for retrospective updates, all four agent models make decisions `as-if' they were always facing a new opponent.

@@

The lexicographic heuristic (algorithm \ref{alg:lexico}) follows the form of that used in \cite{Hau2008}, and assumes a simplified, where an action is a choice between combined lotteries. Functionally, the heuristic maintains a count of the number of times that each action was followed by a payoff, and chooses the action which most commonly has the best payoff, i.e. one reason decision making. This approach requires minimal computation, and does not assume that \(u_{i}\) is static, or known.

@@

%Maybe cut this bit and make explanation richer?
Rather than the psychologically more interesting \ac{PT}, the \ac{CPT}
decision rule is used in this instance, because of the requirement
for women to evaluate more than two `prospects'.\footnote{Where a prospect is a payoff-probability pair, with the set of prospects for an action defining the possible outcomes for it.} \ac{CPT} uses transformed probabilities, underweighting
small probabilities, and overweighting large ones. This is intended to reflect the observed behaviour of humans, where sufficiently high likelihoods are treated as certain, and contrastingly low probabilities as impossible. The correct weighting function is subject to some debate, but here we have used that of \citet{Tversky1992}, which treats probabilities differently for gains and losses -

@@

Where $p$ is the unweighted probability, and $\gamma$ and $\delta$
are the weights for gain and loss probabilities respectively. Along similar lines, the values of losses and gains are transformed to reflect a tendency to regard a loss as more significant than a gain  -

@@

And $\alpha$, and $\beta$ are respectively the power of a gain,
and a loss, and \(x=u_{i}\). The \ac{CPT} value of outcome $x$ is $v(x)w^{+}(x)$
if $x\geq0$, and $v(x)w^{-}(x)$ otherwise. For an action the \ac{CPT}
value is the sum of the value of the prospects of that action, as
in the Bayesian risk model, and the agent chooses the option which maximises this quantity.

### 2014-10-29 Results.tex
Ammendments

\begin{figure}[H]
\begin{adjustbox}{center}\subfloat[Average fraction of population signalling honestly\label{fig:honest_signals}]{\includegraphics[width=0.43\paperwidth]{figures/honesty_plot}

@@

The relationship is not, however, entirely that simple, in that increasing information sharing leads to greater variance between runs. A linear model was was used to predict the between runs interquartile range of the average signal sent by moderate drinkers, with decision rule, and level of information sharing, together with the interaction between the two as predictors. The regression results were significant (\(F_{7,12}=25,p<2.9\times10^{-6}\)) with \(R^2=94\%\), and intercept 0.07. The only significant coefficients were for the interaction terms, which were 0.44 (\(p<0.05\)) for the Bayesian payoff rule, and 0.69 (\(p<0.005\)) for the lexicographic. This suggests that information sharing, for the heuristic style decision rules introduces considerable uncertainty to the model, which is explored further in the sensitivity analysis.

@@

For the \ac{CPT}, and Bayesian decision models, the interaction of bias towards honesty, and distinction between payoffs has a significant, and non-linear effect on instability, and separability of groups. Figure \ref{fig:emulated_payoff_group_iqr} shows the effects, and also highlights the tendency towards poor separability of groups for both the heuristic type decision rules. The response surface of the Bayesian payoff rule is slightly more nuanced than the simple Lexicographic rule, showing better separation close to partial pooling at high payoff distinction, with relatively modest honesty bias.  For the more complex rules, the general tendency is towards less pooling for higher values of both, but with pockets where full pooling occurs.  The plots also suggest that the sensitivity of the \ac{CPT} rule is marginally greater, which reflects the distorted utility function.

>>>>

\begin{figure}[H]
\begin{adjustbox}{center}\subfloat[Average fraction of population ever signalled honestly\label{fig:honest_signals}]{\includegraphics[width=0.43\paperwidth]{figures/honesty_plot}

@@

The relationship is not, however, entirely straighforward, in that increasing information sharing leads to greater variance between runs. A linear model was was used to predict the between runs interquartile range of the average signal sent by moderate drinkers, with decision rule, and level of information sharing, together with the interaction between the two as predictors. The regression results were significant (\(F_{7,12}=25,p<2.9\times10^{-6}\)) with \(R^2=94\%\), and intercept 0.07. The only significant coefficients were for the interaction terms, which were 0.44 (\(p<0.05\)) for the Bayesian payoff rule, and 0.69 (\(p<0.005\)) for the lexicographic. This suggests that information sharing, for the heuristic style decision rules introduces considerable uncertainty to the model, which is explored further in the sensitivity analysis.

@@

For the \ac{CPT}, and Bayesian decision models, the interaction of bias towards honesty, and distinction between payoffs has a significant, and non-linear effect on instability, and separability of groups. Figure \ref{fig:emulated_payoff_group_iqr} shows the effects, and also highlights the tendency towards poor separability of groups for both the heuristic type decision rules. The response surface of the Bayesian payoff rule is slightly more nuanced than the simple Lexicographic rule. Figure \ref{fig:emulated_payoff_group_iqr} shows better separation close to partial pooling at high payoff distinction, with relatively modest honesty bias, which is reflected by the variance contributions of 11\% and 8\% respectively.  For the more complex rules, the general tendency is towards less pooling for higher values of both, but with pockets where full pooling occurs.  The plots also suggest that the sensitivity of the \ac{CPT} rule is marginally greater, which is supported by the significant contribution to variance of close to 15\% for all measures of \(x_{h}\).
